---
title: "House Prices"
author: "Marc Fortó Cornella & Javier Herrer Torres"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
    number_sections: yes
editor_options:
  chunk_output_type: console
subtitle: 'SIM - Assignment 2'
classoption: a4paper
---

GitHub was used as Version Control System for this project.

The contribution of each member is visible through the following repository: https://github.com/marcforto14/Telco_Customer_Churn

And the task distribution: https://github.com/users/marcforto14/projects/1

```
rm(list = ls())
customer_churn <- read.csv("WA_Fn-UseC_-Telco-Customer-Churn.csv")
par(mfrow = c(1, 1))
```

# Data Preparation

## Univariate Descriptive Analysis

### Qualitative and quantitative

No structural errors were found (coding errors, trailing blanks in labels, lower/upper case consistency, etc.).

Data types were propperly adapted to categorial when required (using the `as.factor` function).

Original numeric variables corresponding to qualitative concepts (`SeniorCitizen`) were converted to factors.

```{r}
customer_churn$gender <- as.factor(customer_churn$gender)
customer_churn$SeniorCitizen <- as.factor(
  customer_churn$SeniorCitizen,
  levels = c("No", "Yes")
)
customer_churn$Partner <- as.factor(customer_churn$Partner)
customer_churn$Dependents <- as.factor(customer_churn$Dependents)
customer_churn$PhoneService <- as.factor(customer_churn$PhoneService)
customer_churn$MultipleLines <- as.factor(customer_churn$MultipleLines)
customer_churn$InternetService <- as.factor(customer_churn$InternetService)
customer_churn$OnlineSecurity <- as.factor(customer_churn$OnlineSecurity)
customer_churn$OnlineBackup <- as.factor(customer_churn$OnlineBackup)
customer_churn$DeviceProtection <- as.factor(customer_churn$DeviceProtection)
customer_churn$TechSupport <- as.factor(customer_churn$TechSupport)
customer_churn$StreamingTV <- as.factor(customer_churn$StreamingTV)
customer_churn$StreamingMovies <- as.factor(customer_churn$StreamingMovies)
customer_churn$Contract <- as.factor(customer_churn$Contract)
customer_churn$PaperlessBilling <- as.factor(customer_churn$PaperlessBilling)
customer_churn$PaymentMethod <- as.factor(customer_churn$PaymentMethod)
customer_churn$Churn <- as.factor(customer_churn$Churn)
```


Original numeric varirables corresponding to real quantitative concepts (`MonthlyCharges` and `tenure`) are kept as numeric but additional factors are created as descretization of each numeric variable (`MonthlyChargesGroup` and `TenureGroup`).

```{r}
customer_churn$MonthlyChargesGroup <- cut(
  customer_churn$MonthlyCharges,
  breaks = quantile(
    customer_churn$MonthlyCharges,
    probs = c(0, 0.25, 0.5, 0.75, 1)
  ),
  labels = c("Low", "Medium", "High", "Very high"),
  include.lowest = TRUE
)
customer_churn$TenureGroup <- cut(
  customer_churn$tenure,
  breaks = quantile(
    customer_churn$tenure,
    probs = c(0, 0.25, 0.5, 0.75, 1)
  ),
  labels = c("Very short", "Short", "Long", "Very long"),
  include.lowest = TRUE
)

summary(customer_churn)
```

### Exploratory Data Analysis

The most common number of months which the customer has stayed in the company is very short and the histogram shows a U shape distribution. This is, most clients stay short or long time, and not medium. With clear predominance of the first group (very short tenure).

```{r}
library(ggplot2)

ggplot(customer_churn, aes(x = tenure)) +
  geom_histogram(bins = 20) +
  labs(x = "Tenure (months)", y = "Count", title = "Histogram of tenure")
```

Clients who leave have higher monthly charges than those who don't leave.

```{r}
ggplot(customer_churn, aes(x = Churn, y = MonthlyCharges)) +
  geom_boxplot() +
  labs(
    x = "Churn",
    y = "Monthly charges($)",
    title = "Boxplot of monthly charges by churn"
  )
```

Clients who leave have mostly a month to month contract term.

```{r}
ggplot(customer_churn, aes(x = Contract, fill = Churn)) +
  geom_bar(position = "dodge") +
  labs(
    x = "Contract type",
    y = "Count",
    title = "Barplot of contract type by churn"
  )
```

Clients who leave have are charged a low total amount and high monthly amount.

```{r}
ggplot(
  customer_churn,
  aes(
    x = TotalCharges,
    y = MonthlyCharges,
    color = Churn
  )
) +
  geom_point() +
  labs(
    x = "Total charges ($)",
    y = "Monthly charges ($)",
    title = "Scatterplot of total and monthly charges by churn"
  )
```

Client tenure and monthly charged amount are posivly correlated but not strongly.

```{r}
cor(customer_churn[, c("tenure", "MonthlyCharges")])
```

There is no significant association between gender and churn.

```{r}
chisq.test(customer_churn$gender, customer_churn$Churn)
```
