---
title: "Telco Customer Churn"
author: "Marc Fortó Cornella & Javier Herrer Torres"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
    number_sections: yes
editor_options:
  chunk_output_type: console
subtitle: 'SIM - Assignment 2'
classoption: a4paper
---

GitHub was used as Version Control System for this project.

The contribution of each member is visible through the following repository: https://github.com/marcforto14/Telco_Customer_Churn

And the task distribution: https://github.com/users/marcforto14/projects/1

```{r}
library(car)
library(missMDA)
library(ggplot2)
library(FactoMineR)
library(chemometrics)
library(lmtest)
rm(list = ls())
customer_churn <- read.csv("WA_Fn-UseC_-Telco-Customer-Churn.csv")
par(mfrow = c(1, 1))
```

# Data Preparation

## Univariate Descriptive Analysis

### Qualitative and quantitative

No structural errors were found (coding errors, trailing blanks in labels, lower/upper case consistency, etc.).

Data types were properly adapted to categorical when required (using the `as.factor` function).

Original numeric variables corresponding to qualitative concepts (`SeniorCitizen`) were converted to factors.

```{r}
customer_churn$gender <- as.factor(customer_churn$gender)
customer_churn$SeniorCitizen <- factor(
  customer_churn$SeniorCitizen,
  labels = c("No", "Yes")
)
customer_churn$Partner <- as.factor(customer_churn$Partner)
customer_churn$Dependents <- as.factor(customer_churn$Dependents)
customer_churn$PhoneService <- as.factor(customer_churn$PhoneService)
customer_churn$MultipleLines <- as.factor(customer_churn$MultipleLines)
customer_churn$InternetService <- as.factor(customer_churn$InternetService)
customer_churn$OnlineSecurity <- as.factor(customer_churn$OnlineSecurity)
customer_churn$OnlineBackup <- as.factor(customer_churn$OnlineBackup)
customer_churn$DeviceProtection <- as.factor(customer_churn$DeviceProtection)
customer_churn$TechSupport <- as.factor(customer_churn$TechSupport)
customer_churn$StreamingTV <- as.factor(customer_churn$StreamingTV)
customer_churn$StreamingMovies <- as.factor(customer_churn$StreamingMovies)
customer_churn$Contract <- as.factor(customer_churn$Contract)
customer_churn$PaperlessBilling <- as.factor(customer_churn$PaperlessBilling)
customer_churn$PaymentMethod <- as.factor(customer_churn$PaymentMethod)
customer_churn$Churn <- as.factor(customer_churn$Churn)
```


Original numeric variables corresponding to real quantitative concepts (`MonthlyCharges` and `tenure`) are kept as numeric but additional factors are created as discretization of each numeric variable (`f.MonthlyCharges` and `f.tenure`).

```{r}
customer_churn$f.MonthlyCharges <- cut(
  customer_churn$MonthlyCharges,
  breaks = quantile(
    customer_churn$MonthlyCharges,
    probs = c(0, 0.25, 0.5, 0.75, 1)
  ),
  labels = c("Low", "Medium", "High", "Very high"),
  include.lowest = TRUE
)
customer_churn$f.tenure <- cut(
  customer_churn$tenure,
  breaks = quantile(
    customer_churn$tenure,
    probs = c(0, 0.25, 0.5, 0.75, 1)
  ),
  labels = c("Very short", "Short", "Long", "Very long"),
  include.lowest = TRUE
)
```

### Data quality report

*variable 2: gender*

This is a binary categorical variable with a nearly balanced distribution between females and males. It contains no missing values thus imputation is not needed.
```{r out.width="50%"}
prop.table(table(customer_churn$gender))

# Missing values
sum(is.na(customer_churn$gender))
```

*variable 3: SeniorCitizen*

This is a binary categorical variable. It indicates that 84% of the customers are not senior citizens. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$SeniorCitizen))

# Missing values
sum(is.na(customer_churn$SeniorCitizen))
```

*variable 4: Partner*

This is a binary categorical variable with a nearly balanced distribution between customers that has a partner or not. It contains no missing values thus imputation is not needed.
```{r out.width="50%"}
prop.table(table(customer_churn$Partner))

# Missing values
sum(is.na(customer_churn$Partner))
```

*variable 5: Dependents*

This is a binary categorical variable. It indicates that 70% of the customers have dependents. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$Dependents))

# Missing values
sum(is.na(customer_churn$Dependents))
```

*variable 6: tenure*

tenure is a numerical variable with 0 NA's.  Then we used a Boxplot to visualize the distribution of the values of this variable. The variable has no outliers.
```{r out.width="50%"}
# Missing values
sum(is.na(customer_churn$tenure))
# Univariant Outliers
length(Boxplot(customer_churn$tenure, id = list(n = Inf)))
```

*variable 7: PhoneService*

This is a binary categorical variable. It indicates that 90% of the customers have a phone service. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$PhoneService))

# Missing values
sum(is.na(customer_churn$PhoneService))
```

*variable 8: MultipleLines*

This is a categorical variable with three levels. It indicates that 48% of the customers do not have multiple lines, 42% do have multiple lines, and the remaining 10% do not have phone service. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$MultipleLines))

# Missing values
sum(is.na(customer_churn$MultipleLines))
```

*variable 9: InternetService*

This is a categorical variable with three levels. It indicates that 34% of the customers have DSL, 44% have Fiber optic, and the remaining 22% do not have internet service. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$InternetService))

# Missing values
sum(is.na(customer_churn$InternetService))
```

*variable 10: OnlineSecurity*

This is a categorical variable with three levels. It indicates that 50% of the customers have no online security, 29% have online security, and the remaining 22% do not have internet service. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$OnlineSecurity))

# Missing values
sum(is.na(customer_churn$OnlineSecurity))
```

*variable 11: OnlineBackup*

This is a categorical variable with three levels. It indicates that 44% of the customers have no online backup, 34% have online backup, and the remaining 22% do not have internet service. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$OnlineBackup))

# Missing values
sum(is.na(customer_churn$OnlineBackup))
```

*variable 12: DeviceProtection*

This is a categorical variable with three levels. It indicates that 44% of the customers have no device protection, 34% have device protection, and the remaining 22% do not have internet service. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$DeviceProtection))

# Missing values
sum(is.na(customer_churn$DeviceProtection))
```

*variable 13: TechSupport*

This is a categorical variable with three levels. It indicates that 49% of the customers have no tech support, 29% have tech support, and the remaining 22% do not have internet service. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$TechSupport))

# Missing values
sum(is.na(customer_churn$TechSupport))
```

*variable 14: StreamingTV*

This is a categorical variable with three levels. It indicates that 40% of the customers have no  streaming TV, 38% have streaming TV, and the remaining 22% do not have internet service. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$StreamingTV))

# Missing values
sum(is.na(customer_churn$StreamingTV))
```

*variable 15: StreamingMovies*

This is a categorical variable with three levels. It indicates that 40% of the customers have no streaming movies, 39% have streaming movies, and the remaining 22% do not have internet service. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$StreamingMovies))

# Missing values
sum(is.na(customer_churn$StreamingMovies))
```

*variable 16: Contract*

This is a categorical variable with three levels. It indicates that 55% of the customers have a Month-to-month contract term, 21% have a One year contract term, and the remaining 24% have a Two year contract term. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$Contract))

# Missing values
sum(is.na(customer_churn$Contract))
```

*variable 17: PaperlessBilling*

This is a binary categorical variable. It indicates that 59% of the customers have a paperless billing. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$PaperlessBilling))

# Missing values
sum(is.na(customer_churn$PaperlessBilling))
```

*variable 18: PaymentMethod*

This is a categorical variable with four levels. Electronic check appears to be the most frequently used method, followed by mailed check, credit card (automatic), and bank transfer (automatic) similarly distributed. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$PaymentMethod))

# Missing values
sum(is.na(customer_churn$PaymentMethod))
```

*variable 19: MonthlyCharges*

MonthlyCharges is a numerical variable with 0 NA's. We used a Boxplot to visualize the distribution of the values of this variable. The variable has no outliers.
```{r out.width="50%"}
# Missing values
sum(is.na(customer_churn$MonthlyCharges))
# Univariant Outliers
length(Boxplot(customer_churn$MonthlyCharges, id = list(n = Inf)))
```

*variable 20: TotalCharges*

TotalCharges is a numerical variable with 11 NA's. These NA's correspond to customers with 0 month of tenure with the company. We will then equal their TotalCharges to 0.  We used a Boxplot to visualize the distribution of the values of this variable. The variable has no outliers. We create an additional ordinal TotalCharges factor "f.TotalCharges" to create a discretisation according to the quartiles.
```{r out.width="50%"}
# Missing values
sum(is.na(customer_churn$TotalCharges))
ll_na <- which(is.na(customer_churn$TotalCharges))
#customer_churn[ll_na,]
customer_churn$TotalCharges[ll_na] = 0

# Univariant Outliers
length(Boxplot(customer_churn$TotalCharges, id = list(n = Inf)))

customer_churn$f.TotalCharges <- cut(
  customer_churn$TotalCharges,
  breaks = quantile(
    customer_churn$TotalCharges,
    probs = c(0, 0.25, 0.5, 0.75, 1)
  ),
  labels = c("Low", "Medium", "High", "Very high"),
  include.lowest = TRUE
)
```

*variable 21: Churn*

This is a binary categorical target variable. It indicates that 73% of the customers do not churned. There are no missing values, so imputation is not required.
```{r out.width="50%"}
prop.table(table(customer_churn$Churn))

# Missing values
sum(is.na(customer_churn$Churn))
```

**Inconsistencies**

There are no inconsistencies on the Phone Service and Internet Service variables. 
```{r out.width="50%"}
# Phone Service
sum(customer_churn$PhoneService == "No")
sum(customer_churn$PhoneService == "No" | customer_churn$MultipleLines == "No internet service")
# Internet Service
sum(customer_churn$InternetService == "No")
sum(customer_churn$InternetService == "No" | customer_churn$OnlineSecurity == "No internet service" | customer_churn$OnlineBackup == "No internet service" | customer_churn$DeviceProtection == "No internet service" | customer_churn$TechSupport == "No internet service" | customer_churn$StreamingTV == "No internet service" | customer_churn$StreamingMovies == "No internet service")
```

**Variables**

After exploring all variables, only 'TotalCharges' exhibits missing data, placing it last in terms of missingness. No variables show univariate outliers, suggesting minimal noise in the dataset.

**Multivariate Outliers**

Moutlier is applied on the 3 only numerical variables to find multivariate outliers. A threshold of 2% is chosen as significance level because it already returns a significant amount of outliers, more exactly around 2.5% of instances.
There are 2 types of found outliers:
\begin{itemize}
  \item customers with Very long tenure, Low Monthly Charges and Medium/High Total Charges
  \item customers with Very long tenure, Very high Monthly Charges and Very high Total Charges
\end{itemize}
The first group of customers has only PhoneService, with or without MultipleLines, while the second group has contracted all possible services.
It is chosen not to delete these outliers from the dataset, as they contain valuable information about this type of customers and represent specific customer behaviors.

```{r out.width="50%"}
res.out <- Moutlier(customer_churn[, c(6,19,20)], quantile = 0.98, col = "blue")
#which((res.out$md > res.out$cutoff) & (res.out$rd > res.out$cutoff))
length(which((res.out$md > res.out$cutoff) & (res.out$rd > res.out$cutoff))) / 7043
par(mfrow = c(1, 1))
plot(res.out$md, res.out$rd)
abline(h = res.out$cutoff, col = "red")
abline(v = res.out$cutoff, col = "red")
ll_mvo <- which((res.out$md > res.out$cutoff) & (res.out$rd > res.out$cutoff))
#customer_churn[ll_mvo, ]
```

# Split data

```{r}
set.seed(123)
train_index <- sample(
  seq_len(nrow(customer_churn)),
  size = floor(0.8 * nrow(customer_churn))
)

customer_churn_test_split <- customer_churn[-train_index, ]
customer_churn <- customer_churn[train_index, ]
```

# Exploratory Data Analysis

The most common number of months which the customer has stayed in the company is very short and the histogram shows a U shape distribution. This is, most clients stay short or long time, and not medium. With clear predominance of the first group (very short tenure).

```{r}
library(ggplot2)

ggplot(customer_churn, aes(x = tenure)) +
  geom_histogram(bins = 20) +
  labs(x = "Tenure (months)", y = "Count", title = "Histogram of tenure")
```

Clients who leave have higher monthly charges than those who don't leave.

```{r}
ggplot(customer_churn, aes(x = Churn, y = MonthlyCharges)) +
  geom_boxplot() +
  labs(
    x = "Churn",
    y = "Monthly charges($)",
    title = "Boxplot of monthly charges by churn"
  )
```

Clients who leave have mostly a month to month contract term.

```{r}
ggplot(customer_churn, aes(x = Contract, fill = Churn)) +
  geom_bar(position = "dodge") +
  labs(
    x = "Contract type",
    y = "Count",
    title = "Barplot of contract type by churn"
  )
```

Clients who leave are charged a low Total amount and high monthly amount.
Client tenure and monthly charged amount are positively correlated but not strongly.

```{r}
ggplot(
  customer_churn,
  aes(
    x = TotalCharges,
    y = MonthlyCharges,
    color = Churn
  )
) +
  geom_point() +
  labs(
    x = "Total charges ($)",
    y = "Monthly charges ($)",
    title = "Scatterplot of total and monthly charges by churn"
  )
cor(customer_churn[, c("tenure", "MonthlyCharges")])
```

Customers who leave have mostly an Electronic check payment method.

```{r}
contingency_table <- table(customer_churn$Churn, customer_churn$PaymentMethod);
df_pay_churn <- as.data.frame.matrix(contingency_table)
barplot(t(df_pay_churn), beside = TRUE, legend.text = TRUE, col = c("red", "blue", "green", "purple"), main = "Churn by Payment Method", xlab = "Payment Method", ylab = "Count")
```

There is no significant association between gender and churn.

```{r}
chisq.test(customer_churn$gender, customer_churn$Churn)
```

# Profiling and Feature Selection

The Categorical Data Analysis shows the 5 most important categorical variables to predict the Customer Churn are the Contract, f.tenure (new categorical), OnlineSecurity, TechSupport and InternetService. Diving deep into them, customers who don't leave are associated with a two year contract, long tenure and no internet service. On the other hand, churn is associated with a monthly contract, no online security nor tech support, fiber optic and very short tenure.

Regarding the quantitative variables, MonthlyCharges is the one which contributes the most to customer churn, while TotalCharges and tenure have the opposite effect (negative correlation).

```{r}
catdes_result <- catdes(customer_churn, num.var = 21)

catdes_result$quanti
catdes_result$quanti.var
catdes_result$test.chi2
```

# Churn Modeling

## Numeric variables

From the total charges histogram, the variable is observed as right-skewed, meaning that it has a long tail on the right side of the distribution. This can cause problems for the glm such as violating the assumption of homoscedasticity (the variance of the error term is constant across different values of the predictor variable) affecting the accuracy and reliability of the model estimates. Thus, a log transformation is applied; and a small constant value (0.01) is added avoiding undefined values.

```{r}
hist(
  customer_churn$TotalCharges,
  main = "Histogram of Total Charges",
  xlab = "Total Charges",
  col = "lightblue"
)
```

As seen in class, numeric variables are considered separately applying link functions such as the logit or the orthogonal polynomials where the degree was chosen based on the AIC and BIC criteria avoiding over/underfitting.

```{r}
model_numeric_total_charges <- glm(
  Churn ~ poly(log(TotalCharges + 0.01), 4),
  data = customer_churn,
  family = binomial
)
model_numeric_total_charges
```

```{r}
model_numeric_tenure <- glm(
  Churn ~ poly(tenure, 2),
  data = customer_churn,
  family = binomial
)
model_numeric_tenure
```

```{r}
model_numeric_monthy_charges <- glm(
  Churn ~ poly(MonthlyCharges, 4),
  data = customer_churn,
  family = binomial
)
model_numeric_monthy_charges
```

The additive model results in better AIC and residuals 

```{r}
model_numeric_additive <- glm(
  Churn ~ poly(tenure, 2)
  + poly(MonthlyCharges, 4)
  + poly(log(TotalCharges + 0.01), 4),
  data = customer_churn,
  family = binomial
)
model_numeric_additive
```


The first graph shows how well the model fits the data, and reveals that the model is not linear or homoscedastic, and has some outliers.

The second graph shows how each predictor variable affects the response variable, and reveals that the relationships are nonlinear and have different shapes.

The third graph shows the predicted probability of churn for different values of the predictor variables, and reveals that the interaction effects are not very strong, and the marginal effects are consistent with the second graph.

The fourth graph shows the influence of each observation on the model, and reveals that most of the observations are not influential and fit the model well, but some are influential and do not fit the model well, and are potential outliers.

```{r}
residualPlots(model_numeric_additive)

marginalModelPlots(model_numeric_additive)

library(effects)
plot(allEffects(model_numeric_additive))

influencePlot(model_numeric_additive)
```

## Categorical variables

We initially modify the level "No internet service" to "No" for all internet-related features. This adjustment helps prevent aliasing and collinearity issues among these variables.
We then conduct a VIF analysis on the numerical model and observe excessively high values. As a result, we decide to exclude the variable poly(log(TotalCharges + 0.01), 4) to mitigate collinearity.
Afterwards we added all the categorical variables to our model and created a new model with the output model produced by step.
We then computed the Variance Inflation Factor (VIF), which assesses whether a variable exhibits excessive correlation with other predictors, rendering it redundant in the model. In our analysis, poly(tenure, 2), poly(MonthlyCharges, 4), InternetService, MultipleLines, f.tenure and f.TotalCharges demonstrated relatively high VIF values, indicating a notable level of multicollinearity among these variables. Consequently, we opted to exclude poly(MonthlyCharges, 4), f.tenure and f.TotalCharges from the model due to their significant correlation with other variables. These results lead to a loss of deviance, but there is no apparent collinearity in our model.

```{r}
library(dplyr)
customer_churn <- customer_churn %>%
  mutate(
    OnlineSecurity = recode(OnlineSecurity, "No internet service" = "No"),
    TechSupport = recode(TechSupport, "No internet service" = "No"),
    OnlineBackup = recode(OnlineBackup, "No internet service" = "No"),
    DeviceProtection = recode(DeviceProtection, "No internet service" = "No"),
    StreamingTV = recode(StreamingTV, "No internet service" = "No"),
    StreamingMovies = recode(StreamingMovies, "No internet service" = "No")
  )

vif(model_numeric_additive)
model_numeric_final <- glm(
  Churn ~ poly(tenure, 2)
  + poly(MonthlyCharges, 4),
  data = customer_churn,
  family = binomial
)
vif(model_numeric_final)

m1f <- glm(
  Churn ~ poly(tenure, 2)
  + poly(MonthlyCharges, 4)
  + Contract + OnlineSecurity + TechSupport + InternetService + PaymentMethod 
  + OnlineBackup + MultipleLines + PaperlessBilling + SeniorCitizen + Partner 
  + gender + DeviceProtection + StreamingMovies + StreamingTV + PhoneService 
  + Dependents + f.MonthlyCharges + f.tenure + f.TotalCharges,
  data = customer_churn,
  family = binomial
)

m2f <- step(m1f, trace = 0);m2f
vif(m2f)

# cor(
#   cbind(
#     log(customer_churn$tenure + 0.01),
#     customer_churn$MonthlyCharges,
#     log(customer_churn$TotalCharges + 0.01),
#   )
# )

m3f <- glm(formula = Churn ~ poly(tenure, 2) + Contract + OnlineSecurity + TechSupport + InternetService + 
    PaymentMethod + MultipleLines + PaperlessBilling + SeniorCitizen + 
    StreamingMovies + StreamingTV, 
    family = binomial, data = customer_churn)

anova(m2f, m3f, test = "Chisq")
vif(m3f)
```

There is still a heteroscedasticity problem, variance of the residuals is not constant across different values of the predictor nor the fitted values. The residuals get larger as the prediction moves from small to large or from large to small.

In the influence plot, residuals still range from -2 to 3. However, the Cook's distance has decreased, indicating that the overall fit of the model is more robust, and individual observations have a lesser impact on the results. Some observations with low leverage persist. Nevertheless, these observations demonstrate a low to medium Cook's distance, indicating that they do not exert a substantial influence on the fitted values or estimated coefficients.

```{r}
residualPlots(m3f, layout =c(1,2), ask = FALSE)

influencePlot(m3f)
```

## Main effects and interactions

We begin by applying a log transformation to the poly(tenure, 2) variable, ensuring a constant variance of the error term across different values of the predictor variable. This modification results in a better AIC.

Subsequently, we introduce an interaction between poly(log(tenure + 0.01), 2) and Contract. However, this interaction triggers a warning, suggesting difficulties in logistic regression model convergence, potentially due to perfect separation or quasi-separation in the data. To address this, we opt to eliminate the polynomial term from the tenure variable, creating the model m2i_2 with the interaction between log(tenure + 0.01) and Contract. Comparative analysis between models (m1i and m2i_2) demonstrates a statistically significant improvement, indicating that the interaction term enhances the model's fit.

Moving forward, we incorporate an interaction between SeniorCitizen and PaymentMethod. Similar to the earlier case, the comparison between models (m2i_2 and m3i) reveals a noteworthy improvement, emphasizing the significant contribution of the interaction term to the model.

Lastly, we explore the interaction between InternetService and MultipleLines. Unfortunately, the comparison between models (m3i and m4i) does not yield statistical significance, indicating that this specific interaction does not significantly enhance the model. To finalize our assessment, we compare the models using the AIC. The model with the lowest AIC, m3i in this instance, is deemed the best among the tested models.

```{r}
#poly(log(tenure + 0.01), 2)
m1i <- glm(formula = Churn ~ poly(log(tenure + 0.01), 2) + Contract + OnlineSecurity + TechSupport + InternetService + 
    PaymentMethod + MultipleLines + PaperlessBilling + SeniorCitizen + 
    StreamingMovies + StreamingTV, 
    family = binomial, data = customer_churn)
anova( m3f, m1i, test="Chisq")#significant
AIC(m3f, m1i)
#tenure and contract
m2i <- glm(formula = Churn ~ poly(log(tenure + 0.01), 2) * Contract + OnlineSecurity + TechSupport + InternetService + 
    PaymentMethod + MultipleLines + PaperlessBilling + SeniorCitizen + 
    StreamingMovies + StreamingTV, 
    family = binomial, data = customer_churn)

m2i_2 <- glm(formula = Churn ~ log(tenure + 0.01) * Contract + OnlineSecurity + TechSupport + InternetService + 
    PaymentMethod + MultipleLines + PaperlessBilling + SeniorCitizen + 
    StreamingMovies + StreamingTV, 
    family = binomial, data = customer_churn)

anova( m1i, m2i_2, test="Chisq")#significant
# waldtest( m1i, m2i_2, test="Chisq")

# SeniorCitizen and PaymentMethod
m3i <- glm(formula = Churn ~ log(tenure + 0.01) * Contract + OnlineSecurity + TechSupport + InternetService + 
    MultipleLines + PaperlessBilling + SeniorCitizen * PaymentMethod + 
    StreamingMovies + StreamingTV, 
    family = binomial, data = customer_churn)
anova( m2i_2, m3i, test="Chisq")#significant
waldtest( m2i_2, m3i, test="Chisq" )

# InternetService and MultipleLines
m4i <- glm(formula = Churn ~ log(tenure + 0.01) * Contract + OnlineSecurity + TechSupport + InternetService * MultipleLines 
           + PaperlessBilling + SeniorCitizen * PaymentMethod + StreamingMovies + StreamingTV, family = binomial, data = customer_churn)

anova( m3i, m4i, test="Chisq")#no significant
waldtest( m3i, m4i, test="Chisq" )

AIC(m2i_2, m3i, m4i)
```

We subsequently introduced the polynomial term for tenure, further enhancing our model.

```{r}
m5i <- glm(formula = Churn ~ poly(tenure,2) + log(tenure + 0.01) * Contract + OnlineSecurity + TechSupport + InternetService + MultipleLines 
           + PaperlessBilling + SeniorCitizen * PaymentMethod + StreamingMovies + StreamingTV, family = binomial, data = customer_churn)

anova(m3i, m5i, test="Chisq")#significant
waldtest(m3i, m5i, test="Chisq")
AIC(m3i, m5i)
```

Following the residual analysis, it is evident that a heteroscedasticity issue persists. In the influence plot, residuals still range from -2 to 3, and there has been a slight increase in Cook's distance. However, the latter remains notably low. Additionally, a new observation with higher leverage has emerged. Despite this, the observation exhibits a low Cook's distance, signifying a limited impact on the fitted values or estimated coefficients.

To address these concerns and enhance the model's robustness, we made the decision to exclude three observations with the highest Cook's distance. These particular observations pertain to churn customers with a Two-Year or One-Year Contract, No InternetService, and Medium or Low TotalCharges. This exclusion aims to mitigate the influence of individual observations, thereby contributing to a more resilient model.

```{r}
marginalModelPlots(m5i,id=list(labels=row.names(customer_churn),method=abs(cooks.distance(m5i)), n=5))

residualPlots(m5i, layout=c(2, 2), ask = FALSE)

llnrem<-influencePlot(m5i,id=list(n=3) );llnrem

llistal<-Boxplot(hatvalues(m5i), id.n=10)
abline(h=4*length(coef(m5i))/nrow(customer_churn))
hatvalues(m5i)[llistal]

llistac<-Boxplot(cooks.distance(m5i), id=list(labels=rownames(customer_churn)))
cooks.distance(m5i)[llistac]
customer_churn[llistac,]
llrem <- which( rownames(customer_churn) %in% c("3972","659","5948"))

customer_churn <- customer_churn[-llrem,]

final_model <- glm(formula = Churn ~ poly(tenure,2) + log(tenure + 0.01) * Contract + OnlineSecurity + TechSupport + InternetService + MultipleLines + PaperlessBilling + SeniorCitizen * PaymentMethod + StreamingMovies + StreamingTV, family = binomial, data = customer_churn)
```

# Goodness of fit and Model Interpretation

The McFadden pseudo R-squared has a moderate value which suggests that the model explains some of the variation in the response variable. However, null hypothesis is rejected due to potential collinearity problems. Based on these results, we can say that the model has a moderate predictive power, without a great goodness-of-fit.

```{r}
library(DescTools)
library(ResourceSelection)
library("ROCR")
library(caret)

PseudoR2(final_model, which = "all")

hoslem.test(as.numeric(customer_churn$Churn), fitted(final_model))

model_columns <- c(
  "SeniorCitizen",
  "Dependents",
  "tenure",
  "MultipleLines",
  "InternetService",
  "OnlineSecurity",
  "TechSupport",
  "StreamingTV",
  "StreamingMovies",
  "Contract",
  "PaperlessBilling",
  "PaymentMethod"
)

customer_churn_train <- customer_churn[model_columns]

probabilities_train <- predict(
  final_model,
  type = "response"
)

predictions_train <- ifelse(probabilities_train > 0.5, "Yes", "No")

tt_train <- table(predictions_train, customer_churn$Churn)
tt_train
```

The metrics show that the model correctly classified about 81% of the training data, 67% of the observations that it predicted as positive and 54% of the observations that are actually positive. The F1 score of 60% means that the model has a moderate performance in terms of both precision and recall.

```{r}
accuracy_train <- sum(diag(tt_train)) / sum(tt_train)
accuracy_train
precision_train <- tt_train[2, 2] / (tt_train[2, 2] + tt_train[2, 1])
precision_train
recall_train <- tt_train[2, 2] / (tt_train[2, 2] + tt_train[1, 2])
recall_train
f1_train <- 2 * (
  (precision_train * recall_train)
  / (precision_train + recall_train)
)
f1_train

roc_train <- prediction(
  probabilities_train,
  customer_churn$Churn
)
par(mfrow = c(1, 2))
plot(performance(roc_train, "err"))
plot(performance(roc_train, "tpr", "fpr"))
abline(0, 1, lty = 2)
```

The accuracy and precision metrics are improved in the test data.

```{r}
customer_churn_test <- customer_churn_test_split %>%
  mutate(
    OnlineSecurity = recode(OnlineSecurity, "No internet service" = "No"),
    TechSupport = recode(TechSupport, "No internet service" = "No"),
    OnlineBackup = recode(OnlineBackup, "No internet service" = "No"),
    DeviceProtection = recode(DeviceProtection, "No internet service" = "No"),
    StreamingTV = recode(StreamingTV, "No internet service" = "No"),
    StreamingMovies = recode(StreamingMovies, "No internet service" = "No")
  )

customer_churn_test <- customer_churn_test[model_columns]

probabilities_test <- predict(
  final_model,
  newdata = customer_churn_test
)

predictions_test <- ifelse(probabilities_test > 0.5, "Yes", "No")
table(predictions_test)
table(customer_churn_test$Churn)

tt_test <- table(predictions_test, customer_churn_test_split$Churn)
accuracy_test <- sum(diag(tt_test)) / sum(tt_test)
accuracy_test

precision_test <- tt_test[2, 2] / (tt_test[2, 2] + tt_test[2, 1])
precision_test
recall_test <- tt_test[2, 2] / (tt_test[2, 2] + tt_test[1, 2])
recall_test
f1_test <- 2 * (
  (precision_test * recall_test)
  / (precision_test + recall_test)
)
f1_test

roc_test <- prediction(
  probabilities_test,
  customer_churn_test_split$Churn
)
par(mfrow = c(1, 2))
plot(performance(roc_test, "err"))
plot(performance(roc_test, "tpr", "fpr"))
abline(0, 1, lty = 2)
```